---
title: "MixedModelExample"
author: "James Curley"
date: "July 12, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## An Example Dataset for Mixed Effect Models

These data show `n=60` mice of both sexes. They are from 12 different litters. Their litter/dam id `dam_id` is given as well as two measures of their dam's behavior - licking/grooming `dam_lg` and nursing `dam_nurse`.  The weaning weight of each individual is given `wean_wt` as well as three measures of behavior `of_inner`, `freezing` & `sociability`, and two brain measures `protein1`, `protein2`.

The animals have two grouping variables also.  An `early` treatment which they could either be in the  `control` or `stress` group, and a `post` treatment, which they could either be in the `control` or `treatment` groups.


```{r, message=FALSE, warning=FALSE}
#libraries
library(tidyverse)
```

```{r}
#set seed
set.seed(1)
```


```{r, echo=FALSE}

df <- data.frame(
  id = 1:60,
  dam_id = rep(1:12, each = 5),
  sex = c("M","F"),
  early = rep(c("control", "stress"), each = 30),
  post = sample(c("control", "treatment"), 60, T)
)



#dam lg
lg1 <-  rnorm(n = 6, mean = 8, sd = 2)
lg2 <-  rnorm(n = 6, mean = 3, sd = 1)
lg1[lg1<0]<-0
lg2[lg2<0]<-0
df$dam_lg <- rep(c(lg1, lg2), each = 5)

#dam_nurse
n1 <-  rnorm(n = 6, mean = 70, sd = 8)
n2 <-  rnorm(n = 6, mean = 30, sd = 6)
df$dam_nurse <- rep(c(n1, n2), each = 5)


# wean_wt
df$wean_wt <-
ifelse(df$sex=="M", 
       (log(df$dam_nurse) * 1.025) + runif(n=60, min=5, max=12),
       (log(df$dam_nurse) * 1.005) + runif(n=60, min=4, max=10)
)



# behavior - 1
#sex * lg * early * post
df$of_inner <-
runif(60, 300, 400) + 
  as.numeric(df$sex)*runif(60,40,60) + 
  (df$dam_lg*3.5) + 
  (as.numeric(df$early) * -10 * runif(60,5,15)) +
  (as.numeric(df$post) * -12 * runif(60,1,8))
  
df$of_inner[df$of_inner<0] <- 0 

# behavior - 2 
# post
df$freezing <- 
  runif(60, 0, 20) + 
  (as.numeric(df$post) * 3.5 * runif(60,2.5,2.8))




# behavior - 3
# early * post

df$sociability <-
  runif(60, 260, 320) +
  ((as.numeric(df$early) * -10 * runif(60,5,6.5))) +
  ((as.numeric(df$post) * -12 * runif(60,1,4.8)))






# brain -1 
#sex * wean * early
df$protein1 <-
  round(
    abs(
    runif(60, 0, 4) +
  ((as.numeric(df$sex) * -1 * runif(60,1,1.5))) +
  ((as.numeric(df$early) * -1.2 * runif(60,1,4.8))) +
  ((as.numeric(df$wean) * 0.2 * runif(60,0,1.1)))
  ),1)



# brain - 2
#sex * early * post

df$protein2 <-
  round(
    abs(
      runif(60, 0, 6) +
        ((as.numeric(df$early) * 3.1 * runif(60,1,1.8))) +
        ((as.numeric(df$post) * 1.3 * runif(60,0,3.1)))
    ),1)

```


## Data

```{r}
head(df)

```



## Plots


```{r, message=FALSE, warning=FALSE}

ggplot(df, aes(x=post, y=of_inner, fill=early)) + 
  geom_boxplot()

ggplot(df, aes(x=post, y=freezing, fill=early)) + 
  geom_boxplot()

ggplot(df, aes(x=post, y=sociability, fill=early)) + 
  geom_boxplot()

ggplot(df, aes(x=post, y=protein1, fill=early)) + 
  geom_boxplot()

ggplot(df, aes(x=post, y=protein2, fill=early)) + 
  geom_boxplot()

```
