---
title: "Mixed effects analysis"
author: "Won Lee"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=4, fig.height=3)
```


```{r, message=FALSE, warning=FALSE}
#libraries
library(tidyverse) # I only speak tidyverse 
library(lme4)
library(lmerTest)
library(RColorBrewer)
theme_set(theme_bw()) # set global ggplot theme 
```

These data show `n=60` mice of both sexes. They are from 12 different litters. Their litter/dam id `dam_id` is given as well as two measures of their dam's behavior - licking/grooming `dam_lg` and nursing `dam_nurse`.  The weaning weight of each individual is given `wean_wt` as well as three measures of behavior `of_inner`, `freezing` & `sociability`, and two brain measures `protein1`, `protein2`.

The animals have two grouping variables also.  An `early` treatment which they could either be in the  `control` or `stress` group, and a `post` treatment, which they could either be in the `control` or `treatment` groups.


## Explore the data further 
```{r}
df <- readRDS("df.RDS")

summary(df) 

# notice dam_id and id should be set as character (could be factor), not integer variables.
df <- df %>% 
  mutate_at(vars('id','dam_id'),as.character) 

```

```{r}
table(df$early,df$post)
table(df$early,df$dam_id)
table(df$post,df$dam_id)
table(df$sex,df$dam_id)
```

```{r, fig.width=10}
df %>% 
  gather(y_variables,value,of_inner, freezing, sociability, protein1, protein2) %>% 
  ggplot(aes(x = value))+
  geom_density(alpha = 0.2, size = 1.1)+
  facet_wrap(~y_variables, nrow = 1, scales = "free") # free both x and y axis 
  
```

```{r, fig.height=5,fig.width=10}
df %>% 
  gather(y_variables, value, of_inner, freezing, sociability, protein1, protein2) %>% 
  ggplot(aes(post,value, color = early, fill = early))+
  geom_boxplot(alpha = 0.2, outlier.colour = NA)+
  geom_point(shape = 21, alpha = 0.5, position =  position_dodge(0.75))+
  facet_wrap(~y_variables, nrow = 1, scales = "free_y")+
  scale_color_brewer(palette = 'Dark2')+
  scale_fill_brewer(palette = 'Dark2')

```

```{r}
df %>% 
  gather(y_variables, value, of_inner, freezing, sociability, protein1, protein2) %>% 
  group_by(y_variables, dam_id) %>% 
  summarize_if(is.numeric,var)
```


```{r, fig.height=3, fig.width=10}
df %>% 
  mutate(dam_id = as.numeric(dam_id)) %>% 
  gather(y_variables, value, of_inner, freezing, sociability, protein1, protein2) %>% 
  ggplot(aes(dam_id,value, color = early, fill = early))+
  geom_point(shape = 21, alpha = 0.5, position =  position_dodge(0.75))+
  facet_wrap(~y_variables, nrow = 2, scales = "free_y")+
  scale_x_continuous(breaks =  c(1:12))+
  theme_bw()+
  scale_color_brewer(palette = 'Dark2')+
  scale_fill_brewer(palette = 'Dark2')

```

```{r}
df <- df %>% 
  mutate(my_y_variable = of_inner) #just change here 
```



## Mixed effect model 
- build model
- check model fit
- interpret it 

your_model <- lmer(y_variable ~ x_variable + (1|random_intercept), data = your_data)
```{r}
mod0 <- lmer(my_y_variable ~ 1 + (1|dam_id), data = df, REML = FALSE)
summary(mod0)

mod1 <- lmer(my_y_variable ~ early + (1|dam_id), data = df, REML = FALSE)
summary(mod1)

mod2 <- lmer(my_y_variable ~ post + (1|dam_id), data = df, REML = FALSE)
summary(mod2)

mod3 <- lmer(my_y_variable ~ early + post + (1|dam_id), data = df, REML = FALSE)
summary(mod3)

mod4 <- lmer(my_y_variable ~ early * post + (1|dam_id), data = df, REML = FALSE)
summary(mod4)


AIC(mod0, mod1, mod2, mod3, mod4) # the smaller AIC value is, the better model fit 
```


Want to make things complicated? 
In real experiment examples, 
```{r, fig.width=10}
df %>% 
  gather(y_variables, value, of_inner, freezing, sociability, protein1, protein2) %>% 
  ggplot(aes(post,value, color = interaction(early,sex), fill = interaction(early,sex)))+
  geom_boxplot(alpha = 0.2, outlier.colour = NA)+
  geom_point(shape = 21, alpha = 0.5, position =  position_dodge(0.75))+
  facet_wrap(~y_variables, nrow = 1, scales = "free_y")+
  theme_bw()

```


#### Continuous predictor variables
NOt so good example 
```{r, fig.width=10}
df %>% 
  gather(y_variables, value, of_inner, freezing, sociability, protein1, protein2) %>% 
  ggplot(aes(dam_lg,value))+
  geom_point(shape = 21, alpha = 0.5)+
  geom_smooth(method = "lm")+
  facet_wrap(~y_variables, nrow = 1, scales = "free_y")+
  theme_bw()
```



```{r, fig.width=10}
df %>% 
  gather(y_variables, value, of_inner, freezing, sociability, protein1, protein2) %>% 
  ggplot(aes(dam_nurse,value))+
  geom_point(shape = 21, alpha = 0.5)+
  geom_smooth(method = "lm")+
  facet_wrap(~y_variables, nrow = 1, scales = "free_y")+
  theme_bw()

```

```{r}
mod5 <- lmer(my_y_variable ~ dam_lg + (1|dam_id), data = df, REML = FALSE)
summary(mod5)

mod6 <- lmer(my_y_variable ~ post + dam_lg + (1|dam_id), data = df, REML = FALSE)
summary(mod6)

mod7 <- lmer(my_y_variable ~ post * dam_lg + (1|dam_id), data = df, REML = FALSE)
summary(mod7)

AIC(mod0, mod5, mod6, mod7) # the smaller AIC value is, the better model fit 
```


#### colinearity: Must watch out for this. Again, KNOW YOUR DATA 
```{r}
df %>% 
  ggplot(aes(dam_lg,dam_nurse,label = dam_id, color = early))+
  geom_text()+
  theme_bw()
```

```{r}
# mod1 <- lmer(my_y_variable ~ early + (1|dam_id), data = df, REML = FALSE)
summary(mod1)

mod_do_not <- lmer(my_y_variable ~ early + dam_lg + (1|dam_id), data = df, REML = FALSE)
summary(mod_do_not) # see correlation of fixed effects - 0.9 of correlation 

AIC(mod1,mod_do_not)
```


## some tips 
- extract coefficient 

- multiple comparison: to correct or not to correct??... (Gelman, Hill, Yajima. 2012.)
http://www.stat.columbia.edu/~gelman/research/published/multiple2f.pdf 

